### 2014-06-15 paper/UIST2014-rebuttal.txt

We would first like to thank the reviewers for their many, varied, and thoughtful comments.

CONTRIBUTION CLAIM OVERREACH (R1,R2) - We appreciate that our claims may have sounded overly broad, and agree that instances of interior pipes have been used in certain applications before. We will update statements of contribution to clarify.  For example, on p1, “This paper introduces a novel technique involving removing material...” can be changed to “This paper introduces a novel tool that provides path planning and other high-level functionality to support the design of interior pipes and cavities within 3D models.”

ALGORITHMIC CONTRIBUTION (1AC,R1,R2) - We understand that our system has limitations, and tried to openly describe these in our discussion section. We hope reviewers appreciate that the task of pipe routing is non-trivial and that the work we present offers a contribution to the UIST community; additionally it can serve as a stepping stone to more sophisticated future algorithmic contributions in this area. We can update our discussion section to provide additional insights and possible directions related to the specific limitations mentioned by R1.

We have considered the effectively 2D path-constrained routing problem which R2 brings up; it could be solved using a simulation similar to our endpoint-constrained condition, in which overlapping segments of the connected, Fleury-ized graph are represented as physically-based rods and pushed into the third dimension.  A globally optimum pipe solution requires defining a global pipe energy which can be minimized (based presumably on bending, pipe length, etc.), which we believe is based more on personal preference than global requirements and is outside the scope of this work. However, we will add this discussion to the limitations and future work section.

COMPLEX PIPE GEOMETRIES (R1) - We realize that our focus on useful applications of the technique in the paper does not sufficiently demonstrate the capabilities of our algorithms for routing through objects with complex shapes or topologies. Our tool can indeed route pipes through complex meshes. For example, we can successfully route pipes through knots and around holes in objects (genus >0).  We are happy to add images of routing solutions for more complex geometries in our revision.

SPECULATIVE OBJECTS SECTION (1AC) - In order to strengthen the paper, we are happy to either add additional examples to this section or to remove it, at the discretion of 1AC.

SUPPORT MATERIAL DISCUSSION (1AC,R2) - We agree to shorten this section of the paper and more plainly identify practical difficulties for those who do follow-on work.

VOXELIZATION VS. MESH BOOLEANS (R1) - The use of level sets allowed us more flexibility in, e.g., creating caps on our pipes.  We consider our particular choice orthogonal to the contributions, but this benefit of our voxelization implementation can be made more clear in a final revision.

SUNDRY CORRECTIONS/SUGGESTIONS (R2) - We appreciate this careful list of small changes and will integrate into final submission.

### 2014-06-29 paper/conclusion.tex
making changes thanks to reviews

Pipe mediation opens new interactive device prototyping opportunities both for makers with consumer-grade 3D printers and research and industrial labs with higher-end machines.  While today's 3D printers are not yet able to fabricate active components in-place, we suggest that 3D printed interactive devices can be created with redirection of active input and output via pipes.  This amplifies and extends the utility of existing sensing and actuation approaches.  In addition to describing the design space of pipes, we also presented \systemnamenospace, a design tool for the creation of tubes inside arbitrary 3D models, and discussed many techniques that combine nicely with pipes.  We fabricated several example objects using our tool to explore new points in the design space.  We look forward to extending our software contribution into future multi-material 3D modelling tools.

>>>>

Pipe mediation opens new interactive device prototyping opportunities both for makers with consumer-grade 3D printers and research and industrial labs with higher-end machines.  While today's 3D printers are not yet able to fabricate active components in-place, we suggest that 3D printed interactive devices can be created with redirection of active input and output via pipes.  This amplifies and extends the utility of existing sensing and actuation approaches.  In addition to describing the design space of pipes, we also presented \systemnamenospace, a design tool for the creation of a series of tubes \cite{series-of-tubes} inside arbitrary 3D models, and discussed many techniques that combine nicely with pipes.  We fabricated several example objects using our tool to explore new points in the design space.  We look forward to extending our software contribution into future multi-material 3D modelling tools.

### 2014-06-29 paper/discussion.tex
making changes thanks to reviews

Some media are difficult to insert in pipes, or may not be compatible with all types of pipes.  Threadable materials in particular can provide challenges, as they can get stuck while being fed into pipes. Our system mitagates this by minimizing the bend energy of the interior pipes, but for complex pipe geometries threading could be a challenge.

@@

There are also areas for improvement and future work with our software tool.  The design of \systemname was based on our proposed design space of pipes.  However, the design tool is not yet fully capable of realizing all possibilities in this design space.  Specifically, we did not explicitly explore the mixing or splitting of pipes, and our path-constrained cuts are strictly 2D for now.  A 2.5D SVG import using, e.g., line colors as height would be straightforward, and our described algorithm would continue to work properly.  Furthermore, since pipes are committed to the mesh one at a time, our tool does not currently optimize networks of multiple pipes. Instead, we utilize a greedy algorithm and route a single pipe at a time.  This suggests the need for a better general 3D routing algorithm with multiple traces.  For integration with common electronic components, it would be possible to build a library of ``footprints'' which could be used for recessing them into printed objects' surfaces; such footprints could also be used for automatically generating connecting pipes based on an electrical connection graph, as in circuitboard routing tools like Eagle\footnote{\url{http://www.cadsoftusa.com}}.  Future work should also include assisting makers in the fabrication process itself by automatically cutting objects to reduce support material removal issues. %Additionally, as mentioned there are opportunities to improve the algorithm for neon sign routing, as ours does not guarantee a shortest-path routing.

>>>>

Some media are difficult to insert in pipes, or may not be compatible with all types of pipes.  Threadable materials in particular can provide challenges, as they can get stuck while being fed into pipes. Our system mitigates this by minimizing the bend energy of the interior pipes, but for complex pipe geometries threading could be a challenge.

@@

There are also areas for improvement and future work with our software tool.  The design of \systemname was based on our proposed design space of pipes.  However, the design tool is not yet fully capable of realizing all possibilities in this design space.  Specifically, we did not explicitly explore the mixing or splitting of pipes, and our path-constrained cuts are strictly 2D for now.  A 2.5D SVG import using, e.g., line colors as height would be straightforward, and our described algorithm would continue to work properly.  An extension to \systemname could include detection of where the 2D path intersects itself; intersecting pipe components could be ``pushed into 3D'' using the discrete rods from our endpoint-constrained routing algorithm.  Furthermore, since pipes are committed to the mesh one at a time, our tool does not currently optimize networks of multiple pipes. Instead, we utilize a greedy algorithm and route a single pipe at a time.  This suggests the need for a better general 3D routing algorithm with multiple traces, which could take into account global bending energy, global pipe length, or other task-specific requirements.  For integration with common electronic components, it would be possible to build a library of ``footprints'' which could be used for recessing them into printed objects' surfaces; such footprints could also be used for automatically generating connecting pipes based on an electrical connection graph, as in circuitboard routing tools like Eagle\footnote{\url{http://www.cadsoftusa.com}}.  Future work should also include assisting makers in the fabrication process itself by automatically cutting objects to reduce support material removal issues. %Additionally, as mentioned there are opportunities to improve the algorithm for neon sign routing, as ours does not guarantee a shortest-path routing.

### 2014-06-29 paper/introduction.tex
making changes thanks to reviews

Recently, human-computer interaction researchers have begun to explore how to prototype interactive objects using additive 3D printing processes.
%and maker communities like Thingiverse\footnote{http://thingiverse.com}, recent work has looked at the issue of rapidly prototyping interactive objects. 
Existing approaches include Printed Optics \cite{Willis-printedoptics}, which uses light pipes printed in optically clear material, and Sauron \cite{Savage-sauron}, which uses computer vision to track the movement of components on the inside of a hollow object. These systems exclusively focus on optical sensing and visual displays --- we seek to grow the vocabulary of supported tangible input and output modalities.
%is based on computer vision, but they are limited to visible (reflecting light off fingers, mechanical motion of components) input and visual (screen-like) output; they do not give makers freedom to integrate haptic feedback or other inputs or outputs.

This paper introduces a novel technique involving removing material from the interior of 3D models, prior to printing, to create pipes and other cavities.  These pipes can be filled, post-print, with a variety of media to enable input, display, and tactile feedback.  This {\em subtractive} approach is complementary to {\em additive} approaches that inject custom print materials, such as conductors, during the actual printing process \cite{Sells-reprap}. While our approach requires some manual assembly after a print completes, it gives makers the opportunity to cheaply and rapidly prototype a range of interactive objects on 3D printers, including popular single-material fused-deposition modeling machines (colored objects in Figure \ref{fig:teaser}).

>>>>

Recently, human-computer interaction researchers have begun to explore how to prototype interactive objects using additive 3D printing processes.
Existing approaches, including Printed Optics \cite{Willis-printedoptics} and Sauron \cite{Savage-sauron}, exclusively focus on optical sensing and visual displays --- we seek to grow the vocabulary of supported tangible input and output modalities.

This paper introduces a novel tool that provides path planning and other high-level functionality to support the design of interior pipes and cavities within 3D models.  These pipes can be filled, post-print, with a variety of media to enable input, display, and tactile feedback.  This {\em subtractive} approach is complementary to {\em additive} approaches that inject custom print materials, such as conductors, during the actual printing process \cite{Sells-reprap}. While our approach requires some manual assembly after a print completes, it gives makers the opportunity to cheaply and rapidly prototype a range of interactive objects on 3D printers, including popular single-material fused-deposition modeling machines (colored objects in Figure \ref{fig:teaser}).

### 2014-06-29 paper/relatedwork.tex
making changes thanks to reviews

\subsubsection{Fabrication with Printed Electronics}
Researchers are investigating how to print conductors and place electronic components as part of additive manufacturing.  Sells, et al., \cite{Sells-reprap} use a heated syringe to deposit metal in printed 2D circuitboard channels on objects' surfaces.  Other experiments done by the RepRap project\footnote{\url{http://reprap.org}} use plastic filament blended with conductive materials, however the resistance is currently too high to be useful in circuitry.  Sarik, et al., \cite{Sarik-tracebrush} use aerosol jetting of conductive material onto printed objects.  Park, et al., \cite{Park-microchannels} and Majidi, et al., \cite{Majidi-curvature} inject liquid metals.   Our approach does not deposit such materials automatically, but creates structures inside a model that make manual post-print injection of conductive ink possible.

\subsubsection{Interaction for Fabrication vs. Fabrication for Interaction}
HCI researchers are also developing novel interaction techniques for providing input to digital fabrication machines~\cite{Mueller-constructable,mixfab,Willis-interactive}, and new hybrid hand tools that combine manual processes with digital fabrication~\cite{rivers2012position,zoran2013freed}. These projects are complementary as they target different aspects of the design process, though it is conceivable that our pipe design tools be added to AR modeling interfaces like MixFab~\cite{mixfab}.

>>>>

\subsubsection{Fabrication with Printed Electronics}
Researchers are investigating how to print conductors and place electronic components as part of additive manufacturing.  Sells, et al., \cite{Sells-reprap} use a heated syringe to deposit metal in printed 2D circuitboard channels on objects' surfaces.  Other experiments done by the RepRap project\footnote{\url{http://reprap.org}} use plastic filament blended with conductive materials, however the resistance is currently too high to be useful in circuitry.  Sarik, et al., \cite{Sarik-tracebrush} use aerosol jetting of conductive material onto printed objects.  Park, et al., \cite{Park-microchannels} and Majidi, et al., \cite{Majidi-curvature} inject liquid metals, and Hudson \cite{Hudson-bears} provides techniques for embedding in felted objects.   Our approach does not deposit these materials automatically, but creates structures inside a model that make manual post-print injection of conductive ink possible.

\subsubsection{Interaction for Fabrication vs. Fabrication for Interaction}
HCI researchers are also developing novel interaction techniques for providing input to digital fabrication machines~\cite{Mueller-constructable,mixfab,Willis-interactive}, new applications to extend machine capabilities \cite{Mueller-laserorigami,Hudson-bears}, and new hybrid hand tools that combine manual processes with digital fabrication~\cite{rivers2012position,zoran2013freed}. These projects are complementary as they target different aspects of the design process, though it is conceivable that our pipe design tools be added to AR modeling interfaces like MixFab~\cite{mixfab}.

### 2014-06-29 paper/tubes.tex
making changes thanks to reviews

PBD simulations rarely converge to a steady state, so we simply halt after a fixed number of timesteps (250). Our voxel grid currently uses a resolution of 128x128x128 voxels, which enables interactive operations, but this parameter can be adjusted for different model requirements. The simulation takes about a second for our examples.

@@

%In graph theory, a semi-Eulerian graph is one which has exactly two edges of odd degree.  This kind of graph supports an Euler tour: a path that travels every edge exactly once, beginning and ending at the two nodes of odd degree (in this case, the user-selected start and end nodes).  We create a graph based on the user-specified geometry data, with edges for all line segments; then add edges to connect graph components and make graph the semi-Eulerian.  We find an Euler tour on the modified graph, and thicken the path to create pipes.  
Our technique for routing this path is inspired by work in generating continuous line illustrations from images~\cite{Bosch-tsp,Wong-continuousline}.  Our routing problem is a version of the Chinese Postman Problem\footnote{The CPP is also known as the route inspection problem: \url{http://en.wikipedia.org/wiki/Route_inspection_problem}}, in which we wish to traverse all edges of the graph described by the user's input, much as a postman needs to walk along every road at least once to deliver mail.  In our relaxation, we allow the creation of new edges (i.e., the postman may jump from building to building in addition to using existing roads).
%
%If path components are disconnected, we create edges that connect them; additionally we can create edges connecting odd-degree vertices rather than simply retracing existing edges.  In the final artifact, all created edges will be blocked out by dark material so the inserted medium is not visible (Figure \ref{fig:tool-process-interior}).  
%We offer a short algorithm here, with a more mathematically precise definition and associated proof in supplemental materials appendix.

@@

{\bf Euler Tour:} To find an Euler tour, we use Fleury's algorithm\footnote{\url{http://en.wikipedia.org/wiki/Eulerian_path\#Fleury.27s_algorithm}} with a modification to preserve continuity: at a given node, from multiple candidate edges, we prefer the edge(s) with the smallest angular deviation from the incoming edge (i.e., we prefer to pass straight through a node, if possible).  This minimizes turns in the final artifact, which eases support material removal and assembly.  Our EL wire in Figure \ref{fig:UIST} follows an Euler tour through the input nodes.

@@

To avoid the complexities of direct mesh booleans, we construct a discretized volumetric (i.e., voxel-based) representation of the original mesh and the tube mesh.  Boolean subtraction is trivial in such representations.  We specifically use narrow-band level sets~\cite{Museth04}, which represent the surface as the zero iso-contour of an
approximate distance field.
We then use marching cubes~\cite{Lorensen87} to produce the final mesh for printing.
This strategy does incur some resampling artifacts, however so does the 3D printing; compared to the print process, the time required to voxelize at high resolution is inconsequential.
Since we have a distance field, we can easily introduce useful effects, such as the
ability to add a thin membrane ``cap'' over the end of a tube. 
This is accomplished by applying a standard ``thicken'' operation to the level set version
of the initial mesh, intersecting with a sphere placed at the tube endpoint, and then performing
a boolean union with the initial shape-minus-tube result (Figure \ref{fig:cap}).
A straightforward extension would be to add or subtract additional elements at
the tube endpoints.

@@

\section{Fabrication Techniques}
Several pragmatic concerns arise when fabricating models with hollow pipes and cavities on additive manufactuing machines. Pipes frequently lead to overhangs (where an underlying supporting layer has a smaller footprint than the subsequent layer).
To allow overhangs, 3D printers may lay down support material that must be later removed from the model.  The removal process can be time-intensive and challenging.  The physical fabrication process used by different machines gives rise to differing strategies for avoiding support material or easing its removal.  We next discuss techniques for fused deposition modeling (FDM) machines, such as the Makerbot\footnote{\url{http://makerbot.com}} and polyjet machines such as Stratasys Connex\footnote{\url{http://stratasys.com}}, as well as powder-based machines (e.g., ZCorp\footnote{\url{http://3dsystems.com}}) and those that use stereolithography (e.g., Form 1\footnote{\url{http://formlabs.com}}).

\subsection{Fused Deposition Modeling}
FDM machines lay down lines of extruded plastic filament on top of each other: this allows {\em stepover}, when a layer of plastic is laid slightly offset from the layer below it, and {\em bridging}, when plastic is laid across a gap between two surfaces.  For a layer to fix correctly to its supporting layer, roughly half of the filament needs to overlap.  This means that overhangs without added support material can be approximately $45^{\circ}$, and although cylinders (e.g., pipes) and spheres violate this in some places they can be printed support-free. The colored models in Figure \ref{fig:teaser} were produced on a Makerbot FDM without any support structures inside the pipes.

On single-material hobbyist FDM printers, support structures are built using the same material as the model. Such support may be impossible to remove from internal cavities.  Meshmixer has a built-in tool for minimizing support for FDM prints; it re-orients the part to keep as much surface area as possible within the 45 degree constraint.  In future work, we plan to explore a tool allowing prioritzing specific surfaces (i.e., pipes) for support avoidance, or optimizing pipes to avoid overhangs.

\subsection{Polyjet}
On polyjet machines, bridging is not possible and stepover is more constrained, because the machine is laying down droplets of liquid that are later UV cured.  Unsupported overhangs can be approximately $14^{\circ}$ on our Connex printer.  Polyjet and higher-end FDM machines use a secondary support material that must be manually removed afterwards. Some support can be melted out (Projet wax material); other varieties can be dissolved in lye or blasted with high-pressure water. Complex pipe geometries can make water blasting impossible and other processes more labor- and time-intensive.  All black and clear models in Figure \ref{fig:teaser} required support removal post-print: the neon sign and maze were additionally cut into two pieces along the plane of the pipes to give better access to the support.

\subsection{Powder-based and Stereolithography}
Powder-based and stereolithography (SLA) machines work with a bed of powder or liquid, respectively.  The material at the surface is fused together via glue droplets (for powder) or a UV laser (for SLA).  The remaining powder or liquid serves as support for the model.  These types of printers are perhaps even more suitable for pipe creation than either FDM or multi-jetting printers, as the supporting liquid or loose powder is easier to remove than structural plastic or UV cured material.  We hope to experiment with them in the future.

\subsection{Model Assembly}
For especially complex geometries, models can be ``cut up'' into multiple pieces that can be assembled to avoid support deposition or at least ease material removal; it can also ease insertion of the desired medium. This is somewhat undesirable as it means more time spent in assembly post-print, and additionally a cut pipe may no longer be fluid-tight.  Our plane cut option goes towards this, however a more advanced tool for this purpose is outside the scope of this work.

>>>>

PBD simulations rarely converge to a steady state, so we simply halt after a fixed number of timesteps (250). Our voxel grid currently uses a resolution of 128x128x128 voxels, which enables interactive operations, but this parameter can be adjusted for different model requirements. The simulation takes about a second for our examples, including those in Figure \ref{fig:complex}.

\begin{figure}[h!]
\centering
    \includegraphics[width=3.4in]{figures/routing-tests.png}
\caption{A set of example routings produced by our tool.  Note that because the initial state of the discrete rod is based on an A* routing, it cannot jump around holes in a genus $>0$ mesh (a) even if that would reduce overall bending.  Pipes successfully route through long, curving paths in knots (b), and multiple pipes placed in a single mesh in a tangled configuration successfully avoid each other (c).}
\label{fig:complex}
\end{figure}

@@

%In graph theory, a semi-Eulerian graph is one which has exactly two edges of odd degree.  This kind of graph supports an Euler tour: a path that travels every edge exactly once, beginning and ending at the two nodes of odd degree (in this case, the user-selected start and end nodes).  We create a graph based on the user-specified geometry data, with edges for all line segments; then add edges to connect graph components and make graph the semi-Eulerian.  We find an Euler tour on the modified graph, and thicken the path to create pipes.  
Our technique for routing this path is inspired by work in generating continuous line illustrations from images~\cite{Bosch-tsp,Wong-continuousline}.  Our routing problem is a version of the Chinese Postman Problem~\cite{CPP}, in which we wish to traverse all edges of the graph described by the user's input, much as a postman needs to walk along every road at least once to deliver mail.  In our relaxation, we allow the creation of new edges (i.e., the postman may jump from building to building in addition to using existing roads).
%
%If path components are disconnected, we create edges that connect them; additionally we can create edges connecting odd-degree vertices rather than simply retracing existing edges.  In the final artifact, all created edges will be blocked out by dark material so the inserted medium is not visible (Figure \ref{fig:tool-process-interior}).  
%We offer a short algorithm here, with a more mathematically precise definition and associated proof in supplemental materials appendix.

@@

{\bf Euler Tour:} To find an Euler tour, we use Fleury's algorithm \cite{Fleury} with a modification to preserve continuity: at a given node, from multiple candidate edges, we prefer the edge(s) with the smallest angular deviation from the incoming edge (i.e., we prefer to pass straight through a node, if possible).  This minimizes turns in the final artifact, which eases support material removal and assembly.  Our EL wire in Figure \ref{fig:UIST} follows an Euler tour through the input nodes.

@@

To avoid the complexities of direct mesh booleans and give us additional flexibility, we construct a discretized volumetric (i.e., voxel-based) representation of the original mesh and the tube mesh.  Boolean subtraction is trivial in such representations.  We specifically use narrow-band level sets~\cite{Museth04}, which represent the surface as the zero iso-contour of an
approximate distance field.
We then use marching cubes~\cite{Lorensen87} to produce the final mesh for printing.
This strategy does incur some resampling artifacts, however so does the 3D printing; compared to the print process, the time required to voxelize at high resolution is inconsequential.
Since we have a distance field, we can easily introduce useful effects, such as the
ability to add a thin membrane ``cap'' over the end of a tube. 
This is accomplished by applying a standard ``thicken'' operation to the level set version
of the initial mesh, intersecting with a sphere placed at the tube endpoint, and then performing
a boolean union with the initial shape-minus-tube result (Figure \ref{fig:cap}).
A straightforward extension would be to add or subtract additional elements at
the tube endpoints.

@@

\section{Fabrication Techniques}
Several pragmatic concerns arise when fabricating models with pipes and cavities. Pipes frequently lead to overhangs (where an underlying supporting layer has a smaller footprint than the subsequent layer).
To allow overhangs, 3D printers may lay down support material that must be later removed from the model.  The removal process can be time-intensive and challenging.  The physical fabrication process used by different machines gives rise to differing strategies for avoiding support material or easing its removal.  Techniques differ for fused deposition modeling (FDM) machines, such as the Makerbot\footnote{\url{http://makerbot.com}}, polyjet machines such as Stratasys Connex\footnote{\url{http://stratasys.com}},  powder-based machines (e.g., ZCorp\footnote{\url{http://3dsystems.com}}), and those that use stereolithography (e.g., Form 1\footnote{\url{http://formlabs.com}}).

\subsection{Support Types}
The physical form of support material varies between printer types.  On single-material hobbyist FDM machines, the support is created from the same type of material as the model, laid down in a different pattern to ease detachment.  Multi-material FDM machines typically feature dissolvable support material: by submersing a part in lye, one can eventually remove support material while leaving the model material behind.  Polyjet machines also use a secondary type of material for support: it can be melted out in some cases (e.g., ProJet wax material) or blasted out with high-pressure water.  This blasting process requires fairly open access to the support material: the robot head and maze in Figure \ref{fig:teaser} were cut into two pieces along the plane of the pipes to give better access to the support.  For powder-based and stereolithography machines, the material (powder or photo-cure liquid) that is not used for the model remains in the bed and forms the support structure --- this is uniquely well-suited to removal, as it can simply be poured out after a print is completed.  We hope to experiment with these types of printers more in the future.

\subsection{Model Orientation and Pipe Geometries}
The various types of printers permit different sorts of support-free prints.  FDM allows approximately a $45^\circ$ overhang before support is required.  For polyjet machines, overhangs can be just $14^\circ$.  Reorienting a part within the bed can lead to more or less support required.  MeshMixer already has a built-in tool that reorients models to keep as much surface area as possible within the $45^\circ$ constraint imposed by FDM printers, and alternative pipe cross-section geometries (e.g., teardrop) could alleviate the need for support material to bridge the tops of pipes.  In the future, pipe routes could be optimized to minimize support material required.

\subsection{Model Assembly}
For especially complex geometries, models can be ``cut up'' into multiple pieces that can be assembled, avoiding support deposition or easing material removal; this can also ease insertion of the desired medium. This is somewhat undesirable as it means more time spent in assembly post-print, and additionally a cut pipe may no longer be fluid-tight.  Our plane cut option goes towards this, however a more advanced tool for this purpose is outside the scope of this work.
